maxent model avec xvalidation (year-based test training)

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
april

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava")
source("../src/ipak.R")
ipak(packages)
```

@@

select mont
```{r month of interest -change}
mthna <- "mar"
mthno <- "3"
```

@@

mar2003 b = 1

mar1999 b = 1.25
mar2002 b = 1.25
mar2008 b = 1.25
mar2009 b = 1.25

mar2005 b = 1.5
mar2006 b = 1.5

mar2000 b = 1.75

mar2001 b= 2.25
mar2007 b = 2.25

mar2010 b = 3.5

@@

  if (timeslice == "mar2003"){
    betaval <- 1
  } else if (timeslice == "mar1999"| timeslice == "mar2002" | timeslice == "mar2008" | timeslice == "mar2009"){
    betaval <- 1.25
  } else if (timeslice == "mar2005" | timeslice == "mar2006"){
    betaval <- 1.5
  } else if (timeslice == "mar2000"){
    betaval <- 1.75
  } else if (timeslice == "mar2001" | timeslice == "mar2007"){
    betaval <- 2.25
  } else if (timeslice == "mar2010"){
    betaval <- 3.5
  }

@@

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 2003){
    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    } else if (yr == 2004){
    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    } else if (yr == 2005){
    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    } else if (yr == 2006){
    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
  }
}

aic <- rbind(aic2003, aic2004, aic2005, aic2006)
aicm <- colMeans(aic)
aicm$mod <- "avg"
avgAIC <- rbind(aic, aicm)

write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
```

```{r avg maxres}

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 2003){
    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2004){
    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2005){
    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2006){
    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
  }
}

maxr <- rbind(maxr2003, maxr2004, maxr2005, maxr2006)
maxr$Species <- 1
maxrm <- colMeans(maxr)
maxrm$Species <- "avg"
maxrm <- rbind(maxr, maxrm)

write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
```

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1)
  asclay$V2 <- as.character(asclay$V2)
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclaydep <- asclay$V1

  asc2003 <- raster(asclay$V2)
  asc2004 <- raster(asclay$V3)
  asc2005 <- raster(asclay$V4)
  asc2006 <- raster(asclay$V5)

  stkasc <- stack(asc2003, asc2004, asc2005, asc2006)
  avg <- calc(stkasc, fun = mean)

@@

sum <- calc(stkasc, fun = sum)
avg <- calc(stkasc, fun = mean)
  writeRaster(avg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
  avgplot <- plot(avg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
  dev.off()

>>>>

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava", "raster")
source("../src/ipak.R")
ipak(packages)
```

@@

select mont
```{r month of interest -change}
mthna <- "apr"
mthno <- "4"
```

@@

y1998 = 2.75
y1999 = 2.25
y2000 = 2
y2001 = 1.75
y2002 = 2
y2003 = 2
y2004 = 2
y2005 = 2.25
y2006 = 1.75
y2007 = 1.75
y2008 = 2.25
y2009 = 2.25
y2010 = 2.5
y2011 =  2.5

@@

  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
    betaval <- 1.75
  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
    betaval <- 2
  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
    betaval <- 2.25
  } else if (timeslice == "y2010" | timeslice == "y2011"){
    betaval <- 2.5
  } else if (timeslice == "y1998"){
    betaval <- 2.75
  }

@@

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 1998){
    aic1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1998 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 1999){
    aic1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1999 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2000){
    aic2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2000 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2001){
    aic2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2001 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2002){
    aic2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2002 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2003){
    aic2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2003 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2004){
    aic2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2004 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2005){
    aic2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2005 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2006){
    aic2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2006 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2007){
    aic2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2007 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2008){
    aic2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2008 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2009){
    aic2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2009 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2010){
    aic2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2010 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2011){
    aic2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2011 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2012){
    aic2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2012 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2013){
    aic2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2013 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2014){
    aic2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2014 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2015){
    aic2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2015 <- read.csv(paste0("../output/maxent/models/", timeslice, "/maxentResults.csv"), header = TRUE)

aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)

aicm <- colMeans(aic)
aicm$mod <- "avg"
avgAIC <- rbind(aic, aicm)

maxr$Species <- 1
maxrm <- colMeans(maxr)
maxrm$Species <- "avg"
maxrm <- rbind(maxr, maxrm)

write.csv(avgAIC, paste0("../output/maxent/models/", mthna, "avgAIC.csv"), row.names = FALSE)
write.csv(maxrm, paste0("../output/maxent/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
```

@@

for (x in 1:nrow(depthlayers)){
  asclay <- depthlayers[x, ]
  asclay <- asclay[!is.na(asclay)]
  asclay <- as.data.frame(rbind(asclay))
  asclay$V1 <- as.character(asclay$V1) #v1 corresponds to depth
  asclay$V2 <- as.character(asclay$V2) #v2 to end corresponds to year
  asclay$V3 <- as.character(asclay$V3)
  asclay$V4 <- as.character(asclay$V4)
  asclay$V5 <- as.character(asclay$V5)
  asclay$V6 <- as.character(asclay$V6)
  asclay$V7 <- as.character(asclay$V7)
  asclay$V8 <- as.character(asclay$V8)
  asclay$V9 <- as.character(asclay$V9)
  asclay$V10 <- as.character(asclay$V10)
  asclay$V11 <- as.character(asclay$V11)
  asclay$V12 <- as.character(asclay$V12)
  asclay$V13 <- as.character(asclay$V13)
  asclay$V14 <- as.character(asclay$V14)
  asclay$V15 <- as.character(asclay$V15)

  asclaydep <- asclay$V1

  asc1998 <- raster(asclay$V2)
  asc1999 <- raster(asclay$V3)
  asc2000 <- raster(asclay$V4)
  asc2001 <- raster(asclay$V5)
  asc2002 <- raster(asclay$V6)
  asc2003 <- raster(asclay$V7)
  asc2004 <- raster(asclay$V8)
  asc2005 <- raster(asclay$V9)
  asc2006 <- raster(asclay$V10)
  asc2007 <- raster(asclay$V11)
  asc2008 <- raster(asclay$V12)
  asc2009 <- raster(asclay$V13)
  asc2010 <- raster(asclay$V14)
  asc2011 <- raster(asclay$V15)

  stkasc <- stack(asc1998, asc1999, asc2000, asc2001, asc2002, asc2003, asc2004, asc2005, asc2006, asc2007, asc2008, asc2009, asc2010, asc2011)
  avg <- calc(stkasc, fun = mean)

@@

#stsum <- calc(stkasc, fun=sum)
stavg <- calc(stkasc, fun = mean)
  writeRaster(stavg, (paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".asc")), overwrite=TRUE)
  avgplot <- plot(stavg, png(paste0("../output/maxent/predictions/", mthna, "/avg_", mthna, ".png"))) 
  dev.off()
